---
title: "ML Project"
output: html_document
---


```{r echo=FALSE, warning=FALSE, message=FALSE}
library(caret)
library(ggplot2)
library(plyr)
library(dplyr)
setwd("G:/documentos/Cursos online/DataScience/mlproj")
```

## Load data and preprocess
I load the data with read.csv and since some columns are full of NA I remove them. Finally I also remove timestamps, user_name and the window variables. 

```{r}
data <- read.csv('pml-training.csv', na.strings = c("", "NA"))
nas <- apply(data, 2, function(y) sum(length(which(is.na(y)))))
data <- data[, nas == 0]
data <- select(data, -contains('timestamp'), -X, -user_name, -contains('window'))

test_data <- read.csv('pml-testing.csv', na.strings = c("", "NA"))
test_data <- test_data[, nas == 0]
test_data <- select(test_data, -contains('timestamp'), -X, -user_name, -contains('window'))

```

We are left with `r ncol(data)` predictors and the classe variable.
Then I split the data into training and test set to be able to predict the out of sample error. Instead of using a validation set I will use k-fold cross validation to tune the model.

```{r}
set.seed(1234)
inTrain <- createDataPartition(y=data$classe, p=0.8, list=FALSE)
testing <- data[-inTrain,]
training <- data[inTrain,]
```

I have `r nrow(training)` observations in the training set and `r nrow(testing)` in the testing set.

## Model fitting
Using the caret packet I fit a model using boosting and 5-fold cross validation.
```{r warning=FALSE, message=FALSE}

fitControl <- trainControl(method = "cv", number = 5)
modelFit <- train(classe~., data = training,
                  method = "gbm", trControl = fitControl,
                  verbose = FALSE)
modelFit
```

### Model accuracy
```{r}
ggplot(modelFit)
```


## Model evaluation
Using the part of the data that I reserved for testing I can evaluate the out of sample error more precisely than using the cross validation error:

```{r}
prediction <- predict(modelFit, testing)
confusionMatrix(testing$classe, prediction)
```

From this data we can be confident that the model is very accurate.

## Prediction
Finally I use the model to predict the class fro the 20 observations in the provided test data:

```{r}
answers <- predict(modelFit, test_data)
answers

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```

